---
title: "Data 605 HW 7"
author: "Vishal Arora"
date: "October 12, 2019"
output:
  pdf_document: default
  html_document: default
---

## Q1 :- Let X1, X2, . . . , Xn be n mutually independent random variables, each of which is uniformly distributed on the integers from 1 to k. Let Y denote the minimum of the Xi's. Find the distribution of Y.

Ans :- 
We are given that Y denotes the minimum of the Xis. Suppose each independent random variable Xi has k possibilities.                     

Suppose that each Xi has k possibilities: 1, 2, ..., k. Then, the total possible number of assignments for the entire collection of random variables X1, X2, ..., Xn is (k)^n . This will form the denominator for our probability distribution function.                                       

The number of ways of getting Y = 1 is k^n - (k - 1)^n / k^n, since k^n represents the total number of options and (k-1)^n   represents all of the options where none of the Xi's are equal to 1.                                         

When X = 1:                      
P(X=1) = k^n - (k-1)^n / k^n                      

Similarly When X= 2 & 3:                            
P(X=2) = (k-2+1)^n - (k-2)^n / k^n                                 
P(X=3) = (k-3+1)^n - (k-3)^n / k^n                                       

Generalization this for (X=m):                                                  
P(X=m) = (k-m+1)^n - (k-m)^n * k^n                                      


## Q2:-Your organization owns a copier (future lawyers, etc.) or MRI (future doctors). This machine has a manufacturer's expected lifetime of 10 years. This means that we expect one failure every ten years. (Include the probability statements and R Code for each part.).                                                          


### a) What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a geometric. (Hint: the probability is equivalent to not failing during the first 8 years..)                                

```{r}
p_fail <- 1/10 
n= 8
p_notfail <- 1-p_fail

prob_geom  <- 1-pgeom(n-1, p_fail) 
prob_geom

#Expected value 
expec_val <- 1/p_fail
expec_val

#Standard Deviation
SD <- sqrt(p_notfail/(p_fail^2))
SD 

```

### b) What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as an exponential.                                        

```{r}
n <- 8 
lambda <- 1/10

p_expo <- pexp(n, lambda, lower.tail=FALSE)
round(p_expo,2)

#Expected value = 1/lambda
expec_val <- 1/lambda
expec_val

#SD = 1/??^2
SD <- sqrt(1/lambda^2)
SD 
```

### c) :- What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a binomial. (Hint: 0 success in 8 years)

(n Choose k) p^k * (1-p)^(n???k) 
```{r}
n <- 8
p <- 1/10
q <- (1-p)
k <- 0


p_binomial <- dbinom(k, n, p)
p_binomial


#Expected value
exp_val <- n * p
exp_val


# Standard deviation
sd_bino <- sqrt(n*p*q) 
sd_bino
```

### d) What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a Poisson.

```{r}
#Since average number of failures in every 10 years is 1 so average number of failures in 8 years will be:

lambda <- 8/10
k <- 0

ppois(0,lambda = .8 )

#Expected value 
exp_val <- 8/10
exp_val

# Standard deviation
SD <- sqrt(8/10)
SD
```

