---
title: "Data 605 - Week2_ HW"
author: "Vishal Arora"
date: "September 7, 2019"
output:
  pdf_document: default
  html_document: default
---

# 1. Problem set 1                          

##(a) Show that AT A != A AT in general. (Proof and demonstration.)                             

```{r}

#Let's create a basic 3 x 3 matrix, which we'll call mat1.

mat1 <- matrix(c(1,3,0,4,1,5,2,0,7), nrow = 3, byrow = TRUE)
mat1

# using R get the transpose for mat1 matrices.

mat1T <- t(mat1)
mat1T

#calculate both  ATA and AAT.
ATA <- mat1T %*% mat1
ATA

AAT <- mat1 %*% mat1T
AAT

#It's clear that ATA???AAT for this "general" matrix.
# check programmaticaly AAT == ATA

AAT == ATA


```

## (b) For a special type of square matrix A, we get AT A = AAT . Under what conditions could this be true? (Hint: The Identity matrix I is an example of such a matrix).

```{r}
#Identity Matrix is a square matrix in which all the elements of the principal diagonal are ones and all other elements are zeros.
#Creating Identity matrix using R

A <- diag(3)
A 

#transpose of A
AT <- t(A)
AT

# checking AT A = A AT

(A %*% t(A)) == (t(A) %*% A)


```


# Problem Set 2
## Matrix factorization is a very important problem. There are supercomputers built just to do matrix factorizations. Every second you are on an airplane, matrices are being factorized. Radars that track flights use a technique called Kalman filtering. At the heart of Kalman Filtering is a Matrix Factorization operation. Kalman Filters are solving linear systems of equations when they track your ight using radars.            
## Write an R function to factorize a square matrix A into LU or LDU, whichever you prefer.


```{r}

library(matrixcalc)
matrixFactorization_LU <- function(mat) {
  # Checking whether matrix is a square matrix
 if(is.square.matrix(mat) == FALSE) {
    return("Only square matrix are allowed!")
  }
  
  U <- mat
  n <- dim(mat)[1]
  L <- diag(n)
  
  
  if (n==1) {
    return(list(L,U))
  }
  
  
  for(i in 2:n) {
    for(j in 1:(i-1)) {
      multiplier <- -U[i,j] / U[j,j]
      U[i, ] <- multiplier * U[j, ] + U[i, ]
      L[i,j] <- -multiplier
    }
  }
  return(list(L,U))
}


#using the matrixFactorization_LU
A <- matrix(c(2,0,5,6,1,4,7,3,0), nrow=3, byrow=TRUE)
LU <- matrixFactorization_LU(A)
L<-LU[[1]]  
U<-LU[[2]]

A
L
U

#A == L %*% U
A == L %*% U
```


