---
title: "HW8"
author: "Vishal Arora"
date: "October 19, 2019"
output:
  pdf_document: default
  html_document: default
---

## Company buys 100 lightbulbs, each of which has an exponential lifetime of 1000 hours. What is the expected time for the first of these bulbs to burn out?


```{r}
n  <- 1000
lbulbs <- 100

# The distribution of the minimum value of n independent expontially distributed variables with mean ?? would also be exponential with mean ??(mean)/n(# of items)

b <- n/lbulbs 
b 
```

## Assume that X1 and X2 are independent random variables, each having an exponential density with parameter lambda.  Show that Z = X1-X2 has density .  
$$fZ(z)=(\frac { 1 }{ 2 } )\lambda { e }^{ \lambda -|z| }$$


Ans :-

Both X1 and X2 are evaluated on the interval $$0\le x<\infty.$$  

$$ 1) When\quad X2 \ge X1$$

$$\int _{ 0 }^{ \infty  }{ { fX }^{ (x) }{ fY }^{ x-z }dx }$$
$$\int _{ 0 }^{ \infty  }{ { \lambda e }^{ -\lambda x }{ \lambda e }^{ -\lambda (x???z) } } dx$$
$${\lambda e}^{\lambda z}\int _{ 0 }^{ \infty  }{ { \lambda e }^{ -2\lambda x } } dx$$
$${ \lambda e }^{ \lambda z }(\frac { -1 }{ 2 } { e }^{ -2\lambda x })$$
$$fZ(z)=\frac { \lambda }{ 2 } { e }^{ \lambda z }$$
$$2) When\quad X1 \ge X2$$

$$fZ(z)=\frac { \lambda }{ 2 } { e }^{ -\lambda z }$$
$$fZ(z)=\frac { \lambda }{ 2 } { e }^{ -\lambda |z| }$$
## Let X be a continuous random variable with mean ?? = 10 and variance ??2 = 100/3. Using Chebyshev's Inequality, find an upper bound for the following probabilities.

 pX = var / e^2

 a) P(|X-10|>=2)  
```{r}
var  <- 100/3
e <- 2^2
var/e
```
 
b) P(|X-10|>=5)
```{r}
e<- 5^2
var/e
```
c) P(|X-10|>=9)
```{r}
e <-9^2
var/e
```

d) P(|X-10|>=20)

```{r}
e <- 20^2
var/e
```

