---
title: "HW5"
author: "Vishal Arora"
date: "September 28, 2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# HW5

## Choose independently two numbers B and C at random from the interval [0, 1] with uniform density. Prove that B and C are proper probability distributions.Note that the point (B,C) is then chosen at random in the unit square. 

##Find the probability that                    

(a) B + C < 1/2.                     
(b) BC < 1/2.                                        
(c) |B - C| < 1/2.                                             
(d) max{B,C} < 1/2.                                  
(e) min{B,C} < 1/2.                                    



Q :- Prove that B and C are proper probability distributions
```{r}
# A probability distribution is a statistical function that describes all the possible values and likelihoods that a random variable can take within a given range. This range will be between the minimum and maximum statistically possible values.

n <- 10000
B <- as.data.frame(runif(n, min=0, max=1))
C <- as.data.frame(runif(n, min=0, max=1))



min(B)
max(B)

# checking the skewness of B
hist(B[,1], prob=TRUE, ylim=c(0,1.5), main = 'Histogram of B', xlab = '')


min(C)
max(C)
#checking skewness of C
hist(C[,1], prob=TRUE, ylim=c(0,1.5), main = 'Histogram of B', xlab = '')



j = 0
for(i in 1:length(B)){
  if(B[i]+C[i] < 0.5){
    j = j+1
  }
}
print(paste("The Probabilty B+C < 1/2 =", j/length(B)))


```

As we can see bins of both the histogram are evenly distributed, B & C is proper probability distributions.


(a)
```{r}
a <- sum((B+C) < .5)/n
print(paste("The probability of B+C less than 1/2 is",a))
```

(b)
```{r}
b <- (sum((B*C) < .5)/n)
print(paste("The probability of B*C less than 1/2 is",b))
```

(c) 
```{r}
c <- sum(abs((B-C)) < .5)/n
print(paste("The probability of |B-C| be less than 1/2 is",c))
```

(d)
```{r}
d <- 1
B <- runif(n, min=0, max=1)
C <- runif(n, min=0, max=1)
for(i in 1:length(B)){
  if(max(B[i], C[i]) < 0.5){
    d = d + 1
  }
}
d <- d/10000
print(paste("The probability of max{B,C} less than 1/2 is",d))
```

(e)
```{r}
e <- 1
B <- runif(n, min=0, max=1)
C <- runif(n, min=0, max=1)
for(i in 1:length(B)){
  if(min(B[i], C[i]) < 0.5){
    e = e + 1
  }
}
e <- e/10000
print(paste("The probability of min{B,C} less than 1/2 is",d))
```

